---
title: "Reservoir Gradient"
author: "Jay T. Lennon, Megan L. Larsen, & Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output: 
  pdf_document:
    fig_caption: true
---

Project looking at microbial composition and processes along a reservoir gradient.


# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/ReservoirGradient/analyses")

# Import Required Packages
require("png")
require("ggplot2")      
require("maps")
require("rgdal")
require("raster")
require("OpenMx")
require("reshape")
require("ggmap")
require("grid")
require("akima")
require("plyr")
require("raster") 
require("gridExtra")
```

## FIGURE 1: NUTRIENT PATTERNS ACROSS DAM
```{r}
# Load environmental data
env.dat <- read.csv("../data/ResGrad_EnvDat.csv", header = TRUE)
env.dat <- env.dat[-16,]
```

```{r results='hide', fig.keep='none', warning=FALSE}
# ggplot theme
theme_maps <- function(base_size = 12, base_family = "Arial"){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
          #panel.border = element_rect(color = "black"),
        #panel.margin = unit(1,1,1,1),
        panel.grid.major = element_line(colour = "white"),
        panel.grid.minor = element_line(colour = "white"),
        axis.ticks = element_line(color = "black", size = 1),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 20, face = "bold"),
        legend.position = c(0.9,0.25)
        #,axis.title.x = element_blank(),axis.title.y = element_blank()
    )
}

# get shape files 
## for map inset
usa <- map_data("usa")
IN <- map_data("state", region = "Indiana")

## for main plot
ul <- readOGR("../maps","UniversityLakePoly")
#summary(ul) # Check projection and and datum
#ul <- spTransform(ul, CRS("+proj=longlat +datum=WGS84")) # transform if necessary
ul <- fortify(ul) # raster image for plotting with ggplot2

# If using google map as baselayer
#ul.coords <- c(lon = -86.503087, lat = 39.188686)
#ul.map <- get_map(location = ul.coords, 
#                  zoom = 17, #maptype = "terrain", 
#                  source = "google", color = "bw")
#base.map <- ggmap(ul.map, extent = "device", legend = "topleft")



# Main Map
main.map <- ggplot(aes(long,lat), data = env.dat) + 
  geom_polygon(fill = "grey", size = 1, color = 'black', data = ul, alpha = 1) +
  geom_point(size = 6, shape = 20) +
  theme_maps()  + 
  labs(x = "\nLatitude", y = "Longitude\n")

print(main.map)

# Inset Map
inset <- ggplot() +
  theme_maps() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(color = NULL)) +
  labs(x = NULL, y = NULL)

inset <- inset + geom_map(data = usa, map = usa,
                    aes(x = long, y = lat, map_id = region),
                    color = "black", fill = "#e7e7e7", size = 0.15)

inset <- inset + geom_map(data = IN, map = IN,
                    aes(x = long, y=lat, map_id = region),
                    color = "black", fill = "black", size = 0.15)

inset <- inset + geom_point(aes(x = -86.503087, y = 39.188686), color = "red", size = 2)

print(inset)


# Start Plotting File
png(filename="../figures/Figure1.png",
    width = 1400, height = 1200, res = 96*2)

grid.newpage()
v1 <- viewport(width = 1, height = 1, x = 0.5, y = 0.5) #plot area for the main map
v2 <- viewport(width = 0.4, height = 0.3, x = 0.765, y = 0.31) #plot area for the inset map
print(main.map,vp = v1) 
print(inset,vp = v2)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```


```{r fig.width=6, fig.height=4,echo=FALSE,fig.cap="University Lake Map"}
library(png)
library(grid)
img <- readPNG("../figures/Figure1.png")
grid.raster(img)
```

##FIGURE S1: CHEMICAL PROPERTIES
```{r }

# Start Plotting File
png(filename="../figures/FigureS1.png",
    width = 1200, height = 1200, res = 96*2)

par(mfrow = c(2,2), oma = c(0,2,0,2)+0.5)
    #, oma = c(5, 4, 0, 0) + 0.5)
#par(mfrow = c(1,1), mar = c(1, 1, 1, 7), oma = c(5, 4, 0, 0) + 0.5)

# Total Phosphorus
TP <- plot(env.dat$dist.dam, env.dat$TP, 
     ylab = "", xlab = "", cex.lab = 2, las = 1,
     ylim = c(0,140), xlim = c(-15, 400),
     pch = 22, cex = 2, bg = "white", col = "black", lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0", "40", "80", "120"), at = c(0, 40, 80, 120))
                              
axis(side = 1, lwd.ticks = 2,  labels = F, cex.axis = 2, las = 1, mgp = c(3, 1.5, 0),
   #labels = c("0", "100", "200", "300", "400"), 
   at = c(0, 100, 200, 300, 400))


mtext(expression(paste('Total Phosphorus (',mu,'g P L'^-1*')')), side = 2, line = 4, cex = 1)

# Chlorophyll
chla <- plot(env.dat$dist.dam, env.dat$chla, 
     ylab = "", xlab = "", cex.lab = 2, las = 1,
     ylim = c(0,30), xlim = c(-15, 400),
     pch = 22, cex = 2, bg = "white", col = "black", lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0", "10", "20", "30"), at = c(0, 10, 20, 30))

axis(side = 1, lwd.ticks = 2, labels = F, cex.axis = 2, las = 1, mgp = c(3, 1.5, 0),
   #labels = c("0", "100", "200", "300", "400"), 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('Chlorophyll a (',mu,'g L'^-1*')')), side = 2, line = 4, cex = 1)


#Dissolved Oxygen
plot(env.dat$dist.dam, env.dat$DO, 
     ylab = "", xlab = "", cex.lab = 2, las = 1,
     ylim = c(5,10), xlim = c(-15, 400),
     pch = 22, cex = 2, bg = "white", col = "black", lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("5", "7.5", "10"), at = c(5,7.5, 10))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1, mgp = c(3, 1.5, 0),
   labels = c("0", "100", "200", "300", "400"), 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('Dissolved Oxygen (mg L'^-1*')')), side = 2, line = 4, cex = 1)

#pH
plot(env.dat$dist.dam, env.dat$pH, 
     ylab = "", xlab = "", cex.lab = 2, las = 1,
     ylim = c(8,9), xlim = c(-15, 400),
     pch = 22, cex = 2, bg = "white", col = "black", lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("8", "8.5","9"), at = c(8, 8.5, 9))

axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1, mgp = c(3, 1.5, 0),
   labels = c("0", "100", "200", "300", "400"), 
   at = c(0, 100, 200, 300, 400))

mtext("pH", side = 2, line = 4, cex = 1)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```
## FIGURE 2: METABOLISM ALONG GRADIENT
### MICROBIAL METABOLISM: BP, BR, BGE
### Microbial Functional Groups: Phototroph:Heterotroph
Read in data 
```{r results='hide'}
metab <- read.table("../data/res.grad.metab.txt", sep="\t", header=TRUE)
colnames(metab)[1] <- "dist"
colnames(metab)[2] <- "BP"
colnames(metab)[3] <- "BR"
BGE <- round((metab$BP/(metab$BP + metab$BR)),3)
metab <- cbind(metab, BGE)
```

```{r results='hide'}
png(filename="../figures/Figure2.png",
    width = 1200, height = 1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1, 7, 1, 7), oma = c(5, 4, 0, 0) + 0.5)
bar.layout <- layout(rbind(1, 2, 3), height = c(4, 4, 4)) 
#layout.show(bar.layout)

# Baterial Producivity (BP)

plot(metab$dist, metab$BP, ylab = "", xlab = "", pch = 22, ylim = c(0, 2), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0.0", "1.0", "2.0"), at = c(0, 1.0, 2.0))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 1, 2))

# axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
#     labels = c("0", "100", "200", "300", "400"), at = c(0, 100, 200, 300, 400))

axis(side = 1, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('BP (', mu ,'M C h'^-1* ')')), side = 2, line = 4, cex = 1.5)

# Quadratic regression for BP
dist <- metab$dist
dist2 <- metab$dist^2
BP.fit <- lm(metab$BP ~ dist + dist2)
dist.vals <- seq(0, 375, 25)
BP.pred <- predict(BP.fit,list(dist = dist.vals, dist2 = dist.vals^2))
lines(dist.vals, BP.pred, col = "black", lwd = 2.5, lty = 6)
text(40, 1.8, labels = expression(r^2 == "0.40"), cex = 1.5)

# Bacterial Respiration (BR)

plot(metab$dist, metab$BR, ylab = "", xlab = "", pch = 22, ylim = c(0.75, 3.75), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("1.0", "2.0", "3.0"), at = c(1, 2, 3))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(1, 2, 3))

axis(side = 1, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('BR (', mu ,'M C h'^-1* ')')), side = 2, line = 4, cex = 1.5)

# Simple linear regression for BR
BR.fit <- lm(metab$BR ~ metab$dist)
BR.int <- BR.fit$coefficients[1]
BR.slp <- BR.fit$coefficients[2]
clip(0, 375, 0, 3.75)
abline(a = BR.int, b = BR.slp, col = "black", lwd = 2.5, lty = 6)
text(40, 3.5, labels = expression(r^2 == 0.75), cex = 1.5)

# Bacterial Growth Efficiency

plot(metab$dist, metab$BGE, ylab = "", xlab = "", pch = 22, ylim = c(0, 0.6), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0.0", "0.3", "0.6"), at = c(0, 0.3, 0.6))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 0.3, 0.6))

axis(side = 1, lwd.ticks = 2,  cex.axis = 2, las = 1, mgp = c(3, 1.5, 0),
   labels = c("0", "100", "200", "300", "400"), at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F, cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext("BGE", side = 2, line = 4, cex = 1.5)
mtext("Distance (m)", side = 1, line = 4, cex = 1.5)

# Simple linear regression for BGE
BGE.fit <- lm(metab$BGE ~ metab$dist)
BGE.int <- BGE.fit$coefficients[1]
BGE.slp <- BGE.fit$coefficients[2]
clip(0, 375, 0, 0.58)
abline(a = BGE.int, b = BGE.slp, col = "black", lwd = 2.5, lty = 6)
text(40, 0.535, labels = expression(r^2 == 0.23), cex = 1.5)

# Phototroph to Heterotroph Ratio



dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Microbial Processes"}
library(png)
library(grid)
img <- readPNG("../figures/Figure2.png")
 grid.raster(img)
```


# FIGURE 3: Shifts in Microbial Metabolism
## Crump Model: Mass Effects vs. Species Sorting

### Load required R packages and tools
```{r results = 'hide'}
source("../bin/MothurTools.R")
require("vegan")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```

### Import Shared and Design Files
```{r}
# Define Inputs
# Design = general design file for experiment
# shared = OTU table from mothur with sequence similarity clustering
# Taxonomy = Taxonomic information for each OTU
design <- "../data/UL.design.txt"
shared <- "../data/UL.bac.final.shared"
taxon  <- "../data/UL.bac.final.0.03.taxonomy"

# Import Design
design <- read.delim(design, header=T, row.names=1)

# Import Shared Files
OTUs <- read.otu(shared = shared, cutoff = "0.03")         # 97% Similarity

# Import Taxonomy
OTU.tax <- read.tax(taxonomy = taxon, format = "rdp")
```


# Data Transormations
```{r}
# Remove OTUs with less than two occurences across all sites
OTUs <- OTUs[, which(colSums(OTUs) >= 2)]

# Sequencing an Good's Coverage
# Sequencing Coverage
coverage <- rowSums(OTUs)

# Good's Coverage
goods <- function(x = ""){
  1 - (sum(x == 1) / rowSums(x))
}
goods.c <- goods(OTUs)

# Remove Low Coverage Samples (This code removes two sites: Site 5DNA, Site 6cDNA)
lows <- which(coverage < 10000)
lows
OTUs <- OTUs[-which(coverage < 10000), ]
design <- design[-which(coverage < 10000), ]

# Make Relative Abundence Matrices
OTUsREL <- OTUs
for(i in 1:dim(OTUs)[1]){
  OTUsREL[i,]<- OTUs[i,]/sum(OTUs[i,])
}

# Log Transform Relative Abundances
OTUsREL.log <- decostand(OTUs, method="log")
```

# Calculate Alpha Diversity
```{r}
# Observed Richness
S.obs <- rowSums((OTUs > 0) * 1)

# Simpson's Evenness
SimpE <- function(x = ""){
  x <- as.data.frame(x)
  D <- diversity(x, "inv")
  S <- sum((x > 0) * 1) 
  E <- (D)/S 
  return(E)
}
simpsE <- round(apply(OTUs, 1, SimpE), 3)

# Shannon's Diversity
H <- function(x = ""){
  x <- x[x>0]
  H = 0
  for (n_i in x){
    p = n_i / sum(x)
    H = H - p*log(p) 
  }
  return(H)
}

shan <- round(apply(OTUs, 1, H), 2)
shan2 <- diversity(OTUs, index = "shannon")

alpha.div <- cbind(design, S.obs, simpsE, shan)
```

# Import Phototroph

```{r}
# The phototrophs
cyanos.in <- "../data/UL.cyano.final.shared"
phytos.in <- "../data/UL.euks.final.shared"

cyanos <- read.otu(shared = cyanos.in, cutoff = "0.03")
phytos <- read.otu(shared = phytos.in, cutoff = "0.03")

# Remove OTUs with less than two occurences across all sites
cyanos <- cyanos[, which(colSums(cyanos) >= 2)]
phytos <- phytos[, which(colSums(phytos) >= 2)]

# Remove sites where we have low coverage
cyanos <- cyanos[-which(coverage < 10000), ]
phytos <- phytos[-which(coverage < 10000), ]

# Remove Non Intersecting Sites
ratio.sites <- intersect(intersect(rownames(cyanos), rownames(phytos)), rownames(OTUs))
cyanos <- cyanos[ratio.sites, ]
phytos <- phytos[ratio.sites, ]
heteros <- OTUs[ratio.sites, ]
design.int <- design[ratio.sites, ]

# Remove RNA Sites
DNA.samps <- which(design.int$molecule == "DNA")
cyanos <- cyanos[DNA.samps, ]
phytos <- phytos[DNA.samps, ]
heteros <- OTUs[DNA.samps, ]
design.dna <- design[DNA.samps, ]

# Observed Richness
S.cyano <- rowSums((cyanos > 0) * 1)
S.phyto <- rowSums((phytos > 0) * 1)
S.hetero <- rowSums((heteros > 0) * 1)

N.cyano <- rowSums(cyanos)
N.phyto <- rowSums(phytos)
N.hetero <- rowSums(heteros) - rowSums(cyanos)
                    
HtoC <- N.hetero / N.cyano
HtoP <- N.hetero / N.phyto
HtoBoth <- N.hetero / (N.cyano + N.phyto)


# Ratio Across Gradient Plot
plot(HtoC ~ design.dna$distance, col= "black", pch = 22, las = 1,
     xlim = c(400, 0), ylim = c(0, 1000), cex = 1.5, 
     xlab="", ylab="")

```


## Alpha Diversity Across Gradient
```{r}
# Seperate data based on lake and soil samples
lake <- alpha.div[alpha.div$type == "water",]
soil <- alpha.div[alpha.div$type == "soil", ]

# Calculate Linear Model
model.rich <- lm(lake$S.obs ~ lake$distance * lake$molecule)
summary(model.rich)

# Calculate Confidance Intervals of Model
newdata.rich <- data.frame(cbind(lake$molecule, lake$distance))
conf95.rich <- predict(model.rich, newdata.rich, interval="confidence")

# Average Richess in Terrestrial Habitat
mean(soil$S.obs)

# Dummy Variables Regression Model ("Species Richness"")
D1 <- (lake$molecule == "RNA")*1
fit.Fig.3a <- lm(lake$S.obs ~ lake$distance + D1 + lake$distance*D1)
summary(fit.Fig.3a)

DNA.int.3a <- fit.Fig.3a$coefficients[1]
DNA.slp.3a <- fit.Fig.3a$coefficients[2]
RNA.int.3a <- DNA.int.3a + fit.Fig.3a$coefficients[3]
RNA.slp.3a <- DNA.slp.3a + fit.Fig.3a$coefficients[4]

```

## Similarity To Terrestrial Habitat Across Gradient
```{r}
# Similarity to Soil Sample
UL.bray      <- 1 - as.matrix(vegdist(OTUsREL.log, method="bray"))
UL.bray.lake <- UL.bray[-c(1:3), 1:3] 
bray.mean    <- round(apply(UL.bray.lake, 1, mean), 3)
bray.se      <- round(apply(UL.bray.lake, 1, se), 3)
UL.sim       <- cbind(design[-c(1:3), ], bray.mean, bray.se)

# Calculate Linear Model
model.terr <- lm(UL.sim$bray.mean ~ UL.sim$distance * UL.sim$molecule)
summary(model.terr)

# Calculate Confidance Intervals of Model
newdata.terr <- data.frame(cbind(UL.sim$molecule, UL.sim$distance))
conf95.terr <- predict(model.terr, newdata.terr, interval="confidence")

# Dummy Variables Regression Model ("Terrestrial Influence")
D2 <- (UL.sim$molecule == "RNA")*1
fit.Fig.3b <- lm(UL.sim$bray.mean ~ UL.sim$distance + D2 + UL.sim$distance*D2)
summary(fit.Fig.3b)

DNA.int.3b <- fit.Fig.3b$coefficients[1]
DNA.slp.3b <- fit.Fig.3b$coefficients[2]
RNA.int.3b <- DNA.int.3b + fit.Fig.3b$coefficients[3]
RNA.slp.3b <- DNA.slp.3b + fit.Fig.3b$coefficients[4]
```

## Similarity To Lake Habitat Across Gradient
```{r}
# Similarity to Lake Sample 1
UL.bray2      <- 1 - as.matrix(vegdist(OTUsREL.log, method="bray"))
UL.bray.lake2 <- UL.bray[-c(1:3), 4:7]
UL.sim2       <- cbind(design[-c(1:3), ], "DNA" = apply(UL.bray.lake2[,c(1,3)], 1, mean), 
                                          "RNA" = apply(UL.bray.lake2[,c(2,4)], 1, mean))

# Calculate Linear Model
model.lake1 <- lm(UL.sim2$DNA ~ UL.sim2$distance * UL.sim2$molecule)
model.lake2 <- lm(UL.sim2$RNA ~ UL.sim2$distance * UL.sim2$molecule)
summary(model.lake1)
summary(model.lake2)

# Calculate Confidance Intervals of Model
newdata.lake <- data.frame(cbind(UL.sim2$molecule, UL.sim2$distance))
conf95.lake <- predict(model.lake1, newdata.lake, interval="confidence")

# Dummy Variables Regression Model ("Lake Influence")
D3 <- (UL.sim2$molecule == "RNA")*1
fit.Fig.3c <- lm(UL.sim2$DNA ~ UL.sim2$distance + D3 + UL.sim2$distance*D3)
summary(fit.Fig.3c)

DNA.int.3c <- fit.Fig.3c$coefficients[1]
DNA.slp.3c <- fit.Fig.3c$coefficients[2]
RNA.int.3c <- DNA.int.3c + fit.Fig.3c$coefficients[3]
RNA.slp.3c <- DNA.slp.3c + fit.Fig.3c$coefficients[4]

```

### Figure 3 Plot
```{r}
# Define Plot Parameters
opar <- par()
# par(mar = c(5,6, 1, 1))
mol <- rep(NA, length(lake$molecule))
  for (i in 1:length(mol)){
    if (lake$molecule[i] == "DNA"){
      mol[i] <- 22
    } else {
      mol[i] <- 24
    }
  }
cols <- rep(NA, length(lake$molecule))
  for (i in 1:length(cols)){
    if (lake$molecule[i] == "DNA"){
      cols[i] <- "gray15"
    } else {
      cols[i] <- "gray75"
    }
  }

# Initial Plot
png(filename="../figures/Figure3.png",
    width = 1200, height =1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1, 7, 1, 7), oma = c(5, 4, 0, 0) + 0.5)
bar.layout <- layout(rbind(1, 2, 3), height = c(4, 4, 4)) 

# Richness Across Gradient Plot
plot(lake$S.obs ~ lake$distance, col= "black", bg = cols, pch=mol, las = 1,
     xlim = c(400, 0), ylim = c(0, 2750), cex = 1.5, 
     xlab="", ylab="", xaxt="n")

#   matlines(lake$distance[lake$molecule == "DNA"], conf95.rich[lake$molecule == "DNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))
#   matlines(lake$distance[lake$molecule == "RNA"], conf95.rich[lake$molecule == "RNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))

# Add multiple regression lines
clip(400, 0, 0, 2750)
abline(a = DNA.int.3a, b = DNA.slp.3a, col = "black", lwd = 2.5, lty = 6)

clip(400, 0, 0, 2750)
abline(a = RNA.int.3a, b = RNA.slp.3a, col = "black", lwd = 2.5, lty = 4)

  axis(side = 1, lwd.ticks = 2, labels = F, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)                        
  axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

  # mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Richness\n(S)" , side = 2, line = 4, cex=1.5)

  legend("topright", legend = levels(lake$molecule), pch=c(22, 24), 
       pt.bg = c("gray15", "gray75"), bty='n', cex = 1.25)

  box(lwd=2)

# Terrestrial Influence Plot
#   mol <- rep(NA, length(UL.sim$molecule))
#     for (i in 1:length(mol)){
#       if (UL.sim$molecule[i] == "DNA"){
#         mol[i] <- 21
#       } else {
#         mol[i] <- 24
#       }
#     }
#   cols <- rep(NA, length(UL.sim$molecule))
#     for (i in 1:length(cols)){
#       if (UL.sim$molecule[i] == "DNA"){
#         cols[i] <- "gray15"
#       } else {
#         cols[i] <- "gray75"
#       }
#     }

plot(UL.sim$bray.mean ~ UL.sim$distance, col= "black", bg = cols, pch=mol, las = 1,
     xlim = c(400, 0), ylim = c(0, 0.25), cex = 1.5, 
     xlab="", ylab="", xaxt="n")

#   matlines(lake$distance[lake$molecule == "DNA"], conf95.terr[lake$molecule == "DNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))
#   matlines(lake$distance[lake$molecule == "RNA"], conf95.terr[lake$molecule == "RNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))

# Add multiple regression lines
clip(400, 0, 0, 0.27)
abline(a = DNA.int.3b, b = DNA.slp.3b, col = "black", lwd = 2.5, lty = 6)

clip(400, 0, 0, 0.27)
abline(a = RNA.int.3b, b = RNA.slp.3b, col = "black", lwd = 2.5, lty = 4)

  axis(side = 1, lwd.ticks = 2, labels = F, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=-0.05, labels = F,  cex.axis = 2, las = 1)    
  axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

  # mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Terrestrial\nInfluence" , side = 2, line = 4, cex=1.5)

  legend("topright", legend = levels(UL.sim$molecule), pch=c(22, 24), 
       pt.bg = c("gray15", "gray75"), bty='n', cex = 1.25)

  box(lwd=2)

# Lake Influence Plot
plot(UL.sim2$DNA ~ UL.sim2$distance, col= "black", bg = cols, pch=mol, las = 1,
     xlim = c(400, 0), ylim = c(0, 1), cex = 1.5, 
     xlab="", ylab="")

#   matlines(lake$distance[lake$molecule == "DNA"], conf95.lake[lake$molecule == "DNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))
#   matlines(lake$distance[lake$molecule == "RNA"], conf95.lake[lake$molecule == "RNA", ],
#          lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))

# Add multiple regression lines
clip(400, 0, 0, 1)
abline(a = DNA.int.3c, b = DNA.slp.3c, col = "black", lwd = 2.5, lty = 6)

clip(400, 0, 0, 1)
abline(a = RNA.int.3c, b = RNA.slp.3c, col = "black", lwd = 2.5, lty = 4)

  axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=-0.05, labels = F,  cex.axis = 2, las = 1)  
  axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)           
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)   
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

  mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Lake\nInfluence", side = 2, line = 4, cex=1.5)

  legend("topleft", legend = levels(UL.sim$molecule), pch=c(22, 24), 
       pt.bg = c("gray15", "gray75"), bty='n', cex = 1.25)

  box(lwd=2)



# Close Plot Defice
dev.off()
graphics.off() 
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Microbial Commuinity Shifts"}
library(png)
library(grid)
img <- readPNG("../figures/Figure3.png")
grid.raster(img)
```


# Identifying the Soil Bacteria
```{r}
soil.only <- OTUs[, which(colSums(OTUs[-c(1:3),]) == 0)]
lake.n.soil <- OTUs[, setdiff(colnames(OTUs),colnames(soil.only))]

w.dna <- OTUs[which(design$molecule == "DNA" & design$type == "water"), ]
w.rna <- OTUs[which(design$molecule == "RNA" & design$type == "water"), ]

nvr.act <- which(colSums(w.rna) == 0)

terr.lake <- w.dna[ , c(names(nvr.act))]

terr.rich <- rowSums((terr.lake > 0) * 1)

terr.REL <- rowSums(terr.lake) / rowSums(w.dna) 

design.dna <- design[which(design$molecule == "DNA" & design$type == "water"), ]

terr.rich.log <- log10(terr.rich)
terr.REL.log <- log10(terr.REL)

terr.mod1 <- lm(terr.rich.log ~ design.dna$distance)
terr.mod2 <- lm(terr.REL.log ~ design.dna$distance)

summary(terr.mod1)
summary(terr.mod2)

# # Calculate Confidance Intervals of Model
# newdata.lake <- data.frame(cbind(UL.sim2$molecule, UL.sim2$distance))
# conf95.lake <- predict(terr.mod1, newdata.lake, interval="confidence")
# 
# # Dummy Variables Regression Model ("Lake Influence")
# D3 <- (UL.sim2$molecule == "RNA")*1
# fit.Fig.3c <- lm(UL.sim2$DNA ~ UL.sim2$distance + D3 + UL.sim2$distance*D3)
# summary(fit.Fig.3c)
# 
# DNA.int.3c <- fit.Fig.3c$coefficients[1]
# DNA.slp.3c <- fit.Fig.3c$coefficients[2]
# RNA.int.3c <- DNA.int.3c + fit.Fig.3c$coefficients[3]
# RNA.slp.3c <- DNA.slp.3c + fit.Fig.3c$coefficients[4]

```


# Figure 4: Soil Organisms Plot

```{r}
# Initial Plot
png(filename="../figures/Figure4.png",
    width = 1200, height =1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1, 5, 1, 1), oma = c(4, 4, 0, 0) + 0.5)
bar.layout <- layout(rbind(1, 2), height = c(4, 4)) 

# Soil OTU Richness Across Gradient Plot
plot(terr.rich.log ~ design.dna$distance, col= "black", pch=22, las = 1,
     xlim = c(400, 0), ylim = c(1.5, 3.5), cex = 1.5,  
     xlab="", ylab="", xaxt="n", yaxt="n")

  axis(side = 1, lwd.ticks = 2, labels = T, cex.axis = 1, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 2, lwd.ticks = 2, at = c(2, 3), labels = c(10^2, 10^3), cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, tck = -0.02, at = log10(c(seq(10, 100, by = 10), 
       seq(100, 1000, by = 100), seq(1000, 10000, by = 1000))), labels = F, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, at = c(2, 3), tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 2, lwd.ticks = 2, tck = 0.005, at = log10(c(seq(10, 100, by = 10), 
       seq(100, 1000, by = 100), seq(1000, 10000, by = 1000))), labels = F, cex.axis = 1, las = 1)    
  axis(side = 4, lwd.ticks = 2, at = c(2, 3), tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 4, lwd.ticks = 2, at = c(2, 3), tck=0.02, labels = F,  cex.axis = 2, las = 1)
  axis(side = 4, lwd.ticks = 2, tck = 0.01, at = log10(c(seq(10, 100, by = 10), 
       seq(100, 1000, by = 100), seq(1000, 10000, by = 1000))), labels = F, cex.axis = 1, las = 1)

  # mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Transient\nRichness\n(S)" , side = 2, line = 4, cex=1.5)
  box(lwd=2)
  
#   # Add multiple regression lines
#   clip(400, 0, 0, 1)
#   abline(a = DNA.int.3c, b = DNA.slp.3c, col = "black", lwd = 2.5, lty = 6)
# 
#   clip(400, 0, 0, 1)
#   abline(a = RNA.int.3c, b = RNA.slp.3c, col = "black", lwd = 2.5, lty = 4)
  

# Soil OTU Relative Abundance Across Gradient Plot
plot(terr.REL.log ~ design.dna$distance, col= "black", pch=22, las = 1,
     xlim = c(400, 0), ylim = c(-2.5, -.5), cex = 1.5, 
     xlab="", ylab="", xaxt="n", yaxt="n")

  axis(side = 1, lwd.ticks = 2, labels = T, cex.axis = 1, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 2, lwd.ticks = 2, at = c(-2, -1), labels = c(0.01, 0.1), cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, tck = -0.02, at = log10(c(seq(0.001, 0.01, by = 0.001), 
       seq(0.01, 0.1, by = 0.01), seq(0.1, 1, by = 0.1))), labels = F, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, at = c(-2, -1), tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 2, lwd.ticks = 2, tck = 0.005, at = log10(c(seq(0.001, 0.01, by = 0.001), 
       seq(0.01, 0.1, by = 0.01), seq(0.1, 1, by = 0.1))), labels = F, cex.axis = 1, las = 1)    
  axis(side = 4, lwd.ticks = 2, at = c(-2, -1), tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 4, lwd.ticks = 2, at = c(-2, -1), tck=0.02, labels = F,  cex.axis = 2, las = 1)
  axis(side = 4, lwd.ticks = 2, tck = 0.01, at = log10(c(seq(0.001, 0.01, by = 0.001), 
       seq(0.01, 0.1, by = 0.01), seq(0.1, 1, by = 0.1))), labels = F, cex.axis = 1, las = 1)


  mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Transient\nRelative\nAbundance", side = 2, line = 4, cex=1.5)

  box(lwd=2)  
  

# Close Plot Defice
dev.off()
graphics.off() 
```

```{r fig.width=4, fig.height=4,echo=FALSE,fig.cap="Transient Species Distributions "}
library(png)
library(grid)
img <- readPNG("../figures/Figure4.png")
grid.raster(img)
```

# Define Core Lake Taxa
```{r}
lake.only <- OTUs[, which(colSums(OTUs[c(1:3),]) == 0)]
lake.dna <- lake.only[which(design$molecule == "DNA" & design$type == "water"), ]
lake.rna <- lake.only[which(design$molecule == "RNA" & design$type == "water"), ]

w.dna <- OTUs[which(design$molecule == "DNA" & design$type == "water"), ]
w.rna <- OTUs[which(design$molecule == "RNA" & design$type == "water"), ]

soil.lake.dna <- w.dna[, setdiff(colnames(w.dna), colnames(lake.dna))]
soil.lake.rna <- w.rna[, setdiff(colnames(w.rna), colnames(lake.rna))]

lake.rna.pa <- (lake.rna > 0) * 1
lake.core <- lake.rna[, which((colSums(lake.rna.pa) / nrow(lake.rna.pa)) >= 0.75)]
core.total <- w.dna[, colnames(lake.core)]
core.rich <- rowSums((core.total > 0) * 1)
core.REL <- rowSums(core.total) / rowSums(w.dna) 

```

# Plot Core Community
```{r}
png(filename="../figures/Figure5.png",
    width = 1200, height =1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1, 5, 1, 1), oma = c(4, 4, 0, 0) + 0.5)
bar.layout <- layout(rbind(1, 2), height = c(4, 4)) 

# Soil OTU Richness Across Gradient Plot
plot(core.rich ~ design.dna$distance, col= "black", pch=22, las = 1, 
    ylim = c(12,16), xlim = c(400,0), cex = 1.5, 
     xlab="", ylab="", xaxt="n")

  axis(side = 1, lwd.ticks = 2, labels = F, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)                        
  axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

  # mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Core\nRichness\n(S)" , side = 2, line = 4, cex=1.5)

  box(lwd=2)
  

# Soil OTU Relative Abundance Across Gradient Plot
plot(core.REL ~ design.dna$distance, col= "black", pch=22, las = 1, cex = 1.5, 
     xlab="", ylab="", xlim = c(400, 0))

  axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=-0.05, labels = F,  cex.axis = 2, las = 1)  
  axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)           
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)   
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

  mtext("Distance (m)" , side = 1, line = 3, cex=1.5)
  mtext("Core\nRelative\nAbundance", side = 2, line = 4, cex=1.5)

  box(lwd=2)  


# Close Plot Device
dev.off()
graphics.off() 
```