---
title: "Reservoir Gradient"
author: "Jay T. Lennon, Megan L. Larsen, Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Project looking at microbial composition and processes along a reservoir gradient

# Initial Setup
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/ReservoirGradient")

require("png")
```

# FIGURE 1: NUTRIENT PATTERNS ACROSS DAM
```{r}
require(ggplot2); require(ggmap)
require(reshape); require(plyr)
require(rgdal); require(maptools)
require(raster); require(akima)

# Load environmental data
env.dat <- read.csv("./data/ResGrad_EnvDat.csv", header = TRUE)
env.dat <- env.dat[-16,]

# Import University Lake Polygon
ul <- readOGR("./maps","UniversityLake")
summary(ul) # Check projection and and datum
#ul <- spTransform(ul, CRS("+proj=longlat +datum=WGS84")) # transform if necessary
ul <- fortify(ul) # raster image for plotting with ggplot2

# Download the map from GoogleMaps API - higher quality image than with RgoogleMaps
theMap <- get_map(location = c(lon = -86.503087, lat = 39.188686), 
                  zoom = 17, maptype = "terrain", 
                  source = "google", messaging = F, color = "bw")
base.map <- ggmap(theMap, extent = "device", legend = "topleft")
print(base.map)

# plot map with polygon and sample points
samples <- geom_point(
  aes(x = long, y = lat),
  size = 3, data = env.dat)

ul.pol <- geom_polygon(
  aes(x = long, y = lat, group = group), 
  fill = 'grey', size = 0.5,
  color = 'black', data = ul, alpha=1)

base.map + ul.pol + samples

# Create the gradient contours for the plot
# Add color?

```

# FIGURE 2: METABOLISM ALONG GRADIENT
## MICROBIAL METABOLISM: BP, BR, BGE
## Microbial Functional Groups: Phototroph:Heterotroph
Read in data 
```{r}
metab <- read.table("./data/res.grad.metab.txt", sep="\t", header=TRUE)
colnames(metab)[1] <- "dist"
colnames(metab)[2] <- "BP"
colnames(metab)[3] <- "BR"
BGE <- round((metab$BP/(metab$BP + metab$BR)),3)
metab <- cbind(metab, BGE)
```

```{r}
png(filename="./figures/Figure1.png",
    width = 1200, height = 1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1, 7, 1, 7), oma = c(5, 4, 0, 0) + 0.5)
bar.layout <- layout(rbind(1, 2, 3), height = c(4, 4, 4)) 
#layout.show(bar.layout)

# Baterial Producivity (BP)

plot(metab$dist, metab$BP, ylab = "", xlab = "", pch = 22, ylim = c(0, 2), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0.0", "1.0", "2.0"), at = c(0, 1.0, 2.0))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 1, 2))

# axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
#     labels = c("0", "100", "200", "300", "400"), at = c(0, 100, 200, 300, 400))

axis(side = 1, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('BP (', mu ,'M C h'^-1* ')')), side = 2, line = 4, cex = 1.5)

# Quadratic regression for BP
dist <- metab$dist
dist2 <- metab$dist^2
BP.fit <- lm(metab$BP ~ dist + dist2)
dist.vals <- seq(0, 375, 25)
BP.pred <- predict(BP.fit,list(dist = dist.vals, dist2 = dist.vals^2))
lines(dist.vals, BP.pred, col = "black", lwd = 2.5, lty = 6)
text(40, 1.8, labels = expression(r^2 == "0.40"), cex = 1.5)

# Bacterial Respiration (BR)

plot(metab$dist, metab$BR, ylab = "", xlab = "", pch = 22, ylim = c(0.75, 3.75), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("1.0", "2.0", "3.0"), at = c(1, 2, 3))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(1, 2, 3))

axis(side = 1, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext(expression(paste('BR (', mu ,'M C h'^-1* ')')), side = 2, line = 4, cex = 1.5)

# Simple linear regression for BR
BR.fit <- lm(metab$BR ~ metab$dist)
BR.int <- BR.fit$coefficients[1]
BR.slp <- BR.fit$coefficients[2]
clip(0, 375, 0, 3.75)
abline(a = BR.int, b = BR.slp, col = "black", lwd = 2.5, lty = 6)
text(40, 3.5, labels = expression(r^2 == 0.75), cex = 1.5)

# Bacterial Growth Efficiency

plot(metab$dist, metab$BGE, ylab = "", xlab = "", pch = 22, ylim = c(0, 0.6), xlim = c(-15, 400),
     cex = 2, bg = "white", col = "black", cex.lab = 2, las = 1, lwd = 2,
     yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c("0.0", "0.3", "0.6"), at = c(0, 0.3, 0.6))
                              
axis(side = 4, lwd.ticks = 2, labels = F,  cex.axis = 2, las = 1, 
   at = c(0, 0.3, 0.6))

axis(side = 1, lwd.ticks = 2,  cex.axis = 2, las = 1, mgp = c(3, 1.5, 0),
   labels = c("0", "100", "200", "300", "400"), at = c(0, 100, 200, 300, 400))

axis(side = 3, lwd.ticks = 2, labels = F, cex.axis = 2, las = 1, 
   at = c(0, 100, 200, 300, 400))

mtext("BGE", side = 2, line = 4, cex = 1.5)
mtext("Distance (m)", side = 1, line = 4, cex = 1.5)

# Simple linear regression for BGE
BGE.fit <- lm(metab$BGE ~ metab$dist)
BGE.int <- BGE.fit$coefficients[1]
BGE.slp <- BGE.fit$coefficients[2]
clip(0, 375, 0, 0.58)
abline(a = BGE.int, b = BGE.slp, col = "black", lwd = 2.5, lty = 6)
text(40, 0.535, labels = expression(r^2 == 0.23), cex = 1.5)

# Phototroph to Heterotroph Ratio



dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

# FIGURE 3: Shifts in Microbial Metabolism
## Crump Model: Mass Effects vs. Species Sorting

### Load required R packages and tools
```{r}
source("./bin/MothurTools.R")
require("vegan")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```

### Import Shared and Design Files
```{r}
# Define Inputs
# Design = general design file for experiment
# shared = OTU table from mothur with sequence similarity clustering
# Taxonomy = Taxonomic information for each OTU
design <- "./data/UL.design.txt"
shared <- "./data/UL.bac.final.shared"
taxon  <- "./data/UL.bac.final.0.03.taxonomy"

# Import Design
design <- read.delim(design, header=T, row.names=1)

# Import Shared Files
OTUs <- read.otu(shared = shared, cutoff = "0.03")         # 97% Similarity

# Import Taxonomy
OTU.tax <- read.tax(taxonomy = taxon, format = "rdp")
```

# Data Transormations
```{r}
# Remove OTUs with less than two occurances across all sites
OTUs <- OTUs[, which(colSums(OTUs) >= 2)]

# Sequencing an Good's Coverage
# Sequencing Coverage
coverage <- rowSums(OTUs)
coverage

# Good's Coverage
goods <- function(x = ""){
  1 - (sum(x == 1) / rowSums(x))
}
goods.c <- goods(OTUs)

# Remove Low Coverage Samples (This code removes two sites: Site 5DNA, Site 6cDNA)
lows <- which(coverage < 10000)
lows
OTUs <- OTUs[-which(coverage < 10000), ]
design <- design[-which(coverage < 10000), ]

# Make Relative Abundence Matrices
OTUsREL <- OTUs
for(i in 1:dim(OTUs)[1]){
  OTUsREL[i,]<- OTUs[i,]/sum(OTUs[i,])
}

# Log Transform Relative Abundances
OTUsREL.log <- decostand(OTUs, method="log")

```

# Calculate Alpha Diversity
```{r}
# Observed Richness
S.obs <- rowSums((OTUs > 0) * 1)


# Predicted Richness



# Simpson's Evenness
SimpE <- function(x = ""){
  x <- as.data.frame(x)
  D <- diversity(x, "inv")
  S <- sum((x > 0) * 1) 
  E <- (D)/S 
  return(E)
}
simpsE <- round(apply(OTUs, 1, SimpE), 3)

# Shannon's Diversity
H <- function(x = ""){
  x <- x[x>0]
  H = 0
  for (n_i in x){
    p = n_i / sum(x)
    H = H - p*log(p) 
  }
  return(H)
}

shan <- round(apply(OTUs, 1, H), 2)
diversity(OTUs, index = "shannon")

alpha.div <- cbind(design, S.obs, simpsE, shan)
```


# Alpha Diversity Plots
```{r}
# Seperate data based on lake and soil samples
lake <- alpha.div[alpha.div$type == "water",]
soil <- alpha.div[alpha.div$type == "soil", ]

# Calculate Linear Model
model.1 <- lm(lake$S.obs ~ lake$distance * lake$molecule)
summary(model.1)

# Calculate Confidance Intervals of Model
molecule <- c(rep("DNA", 400), rep("RNA", 400))
distance <- c(rep(seq(1:400), 2))
newdata <- data.frame(cbind(lake$molecule, lake$distance))
conf.95 <- predict(model.1, newdata, interval="confidence")

png(filename="./figures/Figure3.png",
    width = 1200, height = 1200, res = 96*2)


# Richness across Reservoir Gradient
opar <- par()
par(mar = c(5,6, 1, 1))
mol <- rep(NA, length(lake$molecule))
  for (i in 1:length(mol)){
    if (lake$molecule[i] == "DNA"){
      mol[i] <- 21
    } else {
      mol[i] <- 24
    }
  }
cols <- rep(NA, length(lake$molecule))
  for (i in 1:length(cols)){
    if (lake$molecule[i] == "DNA"){
      cols[i] <- "gray15"
    } else {
      cols[i] <- "gray75"
    }
  }

plot(lake$S.obs ~ lake$distance, col= "black", bg = cols, pch=mol, las = 1,
     xlim = c(400, 0), ylim = c(0, 2750), cex = 1.5, 
     xlab="", ylab="")

axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1)
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

mtext("Distance from Dam (m)" , side = 1, line = 3, cex=1.5)
mtext("Observed Richness (S)" , side = 2, line = 4, cex=1.5)

legend("topright", legend = levels(lake$molecule), pch=c(21, 24), 
       pt.bg = c("gray15", "gray75"), bty='n', cex = 1.25)


matlines(lake$distance[lake$molecule == "DNA"], conf.95[lake$molecule == "DNA", ],
         lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))
matlines(lake$distance[lake$molecule == "RNA"], conf.95[lake$molecule == "RNA", ],
         lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))

box(lwd=2)



mean(soil$S.obs)
abline(h = mean(soil$S.obs), lty=2, col="blue")

# Similarity to Soil Sample
UL.bray <- as.matrix(vegdist(OTUsREL.log, method="bray"))
UL.bray.lake <- UL.bray[-c(1:3), 1:3]
bray.mean <- round(apply(UL.bray.lake, 1, mean), 3)
bray.se <- round(apply(UL.bray.lake, 1, se), 3)

UL.sim <- cbind(design[-c(1:3), ], bray.mean, bray.se)


# Calculate Linear Model
model.2 <- lm(UL.sim$bray.mean ~ UL.sim$distance * UL.sim$molecule)
summary(model.2)

# Calculate Confidance Intervals of Model
newdata <- data.frame(cbind(UL.sim$molecule, UL.sim$distance))
conf.95 <- predict(model.2, newdata, interval="confidence")

par(mar = c(5,6, 1, 1))
mol <- rep(NA, length(UL.sim$molecule))
  for (i in 1:length(mol)){
    if (UL.sim$molecule[i] == "DNA"){
      mol[i] <- 21
    } else {
      mol[i] <- 24
    }
  }
cols <- rep(NA, length(UL.sim$molecule))
  for (i in 1:length(cols)){
    if (UL.sim$molecule[i] == "DNA"){
      cols[i] <- "gray15"
    } else {
      cols[i] <- "gray75"
    }
  }

plot(UL.sim$bray.mean ~ UL.sim$distance, col= "black", bg = cols, pch=mol, las = 1,
     xlim = c(400, 0), ylim = c(0.75, 1), cex = 1.5, 
     xlab="", ylab="")

axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1)
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1)
axis(side = 3, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 4, lwd.ticks = 2, tck=-0.01, labels = F,  cex.axis = 2, las = 1)
axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)                              
axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)

mtext("Distance from Dam (m)" , side = 1, line = 3, cex=1.5)
mtext("Bray-Curtis Distance to Soil Samples" , side = 2, line = 4, cex=1.5)

legend("bottomright", legend = levels(UL.sim$molecule), pch=c(21, 24), 
       pt.bg = c("gray15", "gray75"), bty='n', cex = 1.25)


matlines(lake$distance[lake$molecule == "DNA"], conf.95[lake$molecule == "DNA", ],
         lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))
matlines(lake$distance[lake$molecule == "RNA"], conf.95[lake$molecule == "RNA", ],
         lty = c(1, 0, 0), col=c("black", "gray50", "gray50"), lwd=c(2, 1, 1))

box(lwd=2)






dev.off() # this writes plot to folder
graphics.off() # shuts down open devices

```


